---
title: "Cheddar Analysis"
output: html_document
date: "2024-08-30"
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse)
library(readxl)
library(cheddar)
```

Files needed to run code: 
- urban_nodes.csv
- urban_links.csv
- forested_nodes.csv
- forested_links.csv

The study sites are located in Puerto Rico; a tropical, Caribbean island largely composed of mountainous terrain. To conduct the food web analysis, 2 streams with contrasting conditions  were selected.

Using the stream data collected, we will be looking at the:

- Number of Species (S) - NumberOfNodesByClass() 
- Number of Links (L) - NumberOfTrophicLinks()
- Linkage Density (L/S) 
- Connectance (C=L/S^2)

Each node represents one species 

## Río Piedras
Río Piedras is the watershed that drains the San Juan Metropolitan Area; the capital city with a population of approximately 350,000 people.
```{r urban_stream}
dir <- "/Users/reillyconnelly/Desktop/THESIS"
# Manually read the nodes.csv file
urban_nodes <- read.csv("urban_nodes.csv", header = TRUE, stringsAsFactors = FALSE) 
urban_links <- read.csv("urban_links.csv", header = TRUE, stringsAsFactors = FALSE)

urban_web <- Community(
  properties = list(
    title = "Río Piedras Food Web",
    location = "Puerto Rico",     # Overall community location
    ecosystem = "Tropical Rainforest"),
  nodes = urban_nodes ,# Provide nodes directly
  trophic.links = urban_links          # Provide trophic links directly
)

summary(urban_web)
PlotWebByLevel(urban_web)
```

## Quebrada Gatos
Quebrada Gatos is a pristine stream located in El Yunque National Forest; a tropical rainforest in the Northeast region.
```{r forested_stream}
forested_nodes <- read.csv("forested_nodes.csv", header = TRUE, stringsAsFactors = FALSE) 
forested_links <- read.csv("forested_links.csv", header = TRUE, stringsAsFactors = FALSE)

forested_web <- Community(
  properties = list(
    title = "Quebrada Gatos",
    location = "Puerto Rico",     # Overall community location
    ecosystem = "Tropical Rainforest"),
  nodes = forested_nodes ,# Provide nodes directly
  trophic.links = forested_links          # Provide trophic links directly
)

summary(forested_web)
PlotWebByLevel(forested_web)
```

We observe less diversity in forested streams because of primary production like algae. Urban streams in PR have a lot of nutrients that stimulate algae, causing diversity and more complexity. In contrast forested streams have a more closed canopy and less primary production resulting in comparatively less diversity. (i.e. less taxa, less complex, less species, high connectance) 
