---
title: "Comparative Analysis"
output: html_document
date: "2025-03-31"
---

```{r setup}
knitr::opts_chunk$set(echo = FALSE)
library(readxl)
library(dplyr)
library(ggplot2)
library(patchwork)
library(broom)
library(car)

data <- read_xlsx('data.xlsx', sheet='final_data')

# Exclude the last row (Yule et al. 2010) because data includes algae as species richness which influences the metrics
# Results significant with and without it 
data <- data[-nrow(data), ]
data <- data %>%
  mutate(log_num_spec = log(num_spec),
         log_num_link = log(num_link), 
         log_connect = log(connect))
head(data)
```

```{r residuals, echo=TRUE}
# Create a new column to mark "our study"
data$highlight <- ifelse(data$paper == "our study", "Our Study", "Other Studies")

# Fit the linear model
model <- lm(log_connect ~ log_num_spec, data = data)
data$residuals <- resid(model)

# Plot residuals with highlighted points
ggplot(data, aes(x = region, y = residuals, fill = region)) + 
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +  # Boxplot without showing default outliers
  geom_jitter(aes(color = highlight), width = 0.2, alpha = 0.7, size = 3) +  # Jittered points
  scale_fill_manual(values = c("tropical" = "#E69F00", "temperate" = "#0072B2")) +  
  scale_color_manual(values = c("Our Study" = "red", "Other Studies" = "black")) +  # Custom colors
  labs(x = "Region", y = "Residuals", title = "Residual Comparison of Food Web Connectance",
       color = "Study Type") +
  theme_minimal() +
  theme(legend.position = "top")

```

We use ANCOVA because it allows us to compare the slopes of the relationship between species richness and connectance while controlling for the effect of region. Unlike simply comparing individual regression slopes, ANCOVA accounts for both the differences in the slopes and the intercepts between the regions. This test combines the effects of both the categorical variable (region) and the continuous variable (species richness) in a single model, providing a more robust and accurate comparison. By doing this, ANCOVA helps us determine if the relationship between species richness and connectance differs significantly between regions, while controlling for the influence of region itself.

## Number of Links vs. Species Richness

This graph models the **linear regression line** predicting log_num_links by log_num_spec **within each region**, just based on the **raw data**

```{r species vs links}
# ANCOVA
mod_links <- lm(log_num_link ~ log_num_spec * region, data = data)
summary(mod_links)

# Create the plot with linear models for both regions on the same plot
a <- ggplot(data, aes(x = log_num_spec, y = log_num_link, 
                      color = region)) +
  geom_point(size = 3) +  # Points colored by region
  geom_smooth(method = "lm", se = T) +  # Add separate linear models for each region
  labs(
    title = "Links Tropical vs. Temperate Regions",
    x = "Log Number of Species",
    y = "Log Number of Links",
    color = "Region"
  ) +
  theme_minimal() +
  theme(legend.position = "top")  # Move the legend to the top
a
```

```{r}
# diagnostics on model
data_aug <- augment(mod_links)

# Plot Cook's distance
ggplot(data_aug, aes(x = .cooksd)) +
  geom_histogram(bins = 30, fill = "tomato") +
  labs(title = "Cook's Distance for Influential Points")

# View top influential points
data_aug %>%
  arrange(desc(.cooksd)) %>%
  head(10) 

data %>%
  filter(num_spec > 75) %>%
  select(num_spec, num_link, region, paper) %>%
  arrange(desc(num_spec))

library(ggrepel)

ggplot(data, aes(x = num_spec, y = num_link, color = region)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = TRUE) +
  geom_text_repel(
    data = data %>% filter(num_spec > 75),
    aes(label = paper), 
    size = 3, max.overlaps = 10
  ) +
  labs(title = "High-Species Food Webs by Source Paper",
       x = "Number of Species", y = "Number of Links") +
  theme_minimal()
```

```{r}
low_spec <- data |> 
  filter(num_spec < 75) |> 
  select(log_num_spec, log_num_link, region, paper)

low_spec_mod <- lm(log_num_link ~ log_num_spec * region, data = low_spec)
summary(mod_links)

ggplot(low_spec, aes(x = num_spec, y = num_link, 
                      color = region)) +
  geom_point(size = 3) +  # Points colored by region
  geom_smooth(method = "lm", se = T) +  # Add separate linear models for each region
  labs(
    title = "Links Tropical vs. Temperate Regions",
    x = "Number of Species",
    y = "Number of Links",
    color = "Region"
  ) +
  theme_minimal() +
  theme(legend.position = "top") 
```

The significant interaction term (p = 0.000566) shows that species richness effects the number of links differently across regions (i.e. the effect of region on num_link). Temperate streams have a steeper increase in links (slope = 1.7022), while tropical streams have a slower increase (slope = 1.1655) but start with more links. This suggests temperate streams form interactions more flexibly, whereas tropical streams may have more constrained or specialized networks.

Summary of the Slopes:

-   Temperate Region:

    -   Slope = log_num_spec + coefficient = 1.7022

    -   In temperate regions, for every 1-unit increase in log number of species, log number of links increases by 1.702.

-   Tropical Region:

    -   Slope = 1.7022 + (-0.5367) = 1.1655

    -   In tropical regions, for every 1-unit increase in log number of species, log number of links increases by 1.1655

## Species Richness--Connectance Relationship

This graph models the **linear regression line** predicting log_connect by log_num_spec **within each region**, just based on the **raw data**

```{r species vs connectance}
# ANCOVA
mod_connect <- lm(log_connect ~ log_num_spec * region, data = data)
summary(mod_connect)

# Plot log-log relationship with regression lines and confidence intervals
ggplot(data, aes(x = log_num_spec, y = log_connect, color = region)) +
  geom_point(alpha = 0.6, size = 3) +  # Add points with transparency
  geom_smooth(method = "lm", se = TRUE, fullrange = TRUE, alpha = 0.2) +  # Regression lines with confidence intervals
  scale_color_manual(values = c("tropical" = "#E69F00", "temperate" = "#0072B2")) +  # Custom colors
  labs(x = "Log Number of Species", 
       y = "Log Connectance", 
       title = "Connectance in Tropical vs. Temperate Regions",
       color = "Region") +
  theme_minimal() +
  theme(legend.position = "top", text = element_text(size = 14))

# Extracting Slopes from the ANCOVA Model:
# Extracting coefficients from the ANCOVA model
coefficients_ancova <- coef(mod_connect)

# Slope for temperate region
slope_temperate <- coefficients_ancova["log_num_spec"]
cat("Slope for Temperate Region:", slope_temperate, "\n")

# Slope for tropical region (sum of log_num_spec and log_num_spec:regiontropical)
slope_tropical <- coefficients_ancova["log_num_spec"] + coefficients_ancova["log_num_spec:regiontropical"]
cat("Slope for Tropical Region:", slope_tropical, "\n")
```

The significant interaction term (p = 0.000642) shows that the relationship between species richness and connectance is different across regions (i.e. the effect of region on log_num_spec). Tropical streams have a steeper decrease in log connectance (slope = -0.8343), while tropical streams have a slower decrease (slope = -0.3014).

Summary of the Slopes:

-   Temperate Region:

    -   Slope = -0.3014

    -   In temperate regions, for every 1-unit increase in log number of species, log connectance decreases by 0.3014.

-   Tropical Region:

    -   Slope = -0.3014 + (-0.53294) = -0.8343

    -   In tropical regions, for every 1-unit increase in log number of species, log connectance decreases by 0.8343.

```{r nls}
# Create the plot with linear models for both regions on the same plot
# Fit an inverse model for each region
  model_results <- data %>%
  group_by(region) %>%
  do(model = nls(num_spec ~ a / connect + b, data = ., start = list(a = 1, b = 1)))

# Generate predicted values for plotting
predicted_data <- data %>%
  group_by(region) %>%
  do({
    model <- nls(connect ~ a / num_spec + b, data = ., start = list(a = 1, b = 0))
    data.frame(num_spec = seq(min(.$num_spec), max(.$num_spec), length.out = 100),
               connect = predict(model, newdata = data.frame(num_spec = seq(min(.$num_spec), max(.$num_spec), length.out = 100))),
               region = unique(.$region))
  })

# Plot the data and fitted curves
b <- ggplot(data, aes(x = num_spec, y = connect, color = region)) +
  geom_point(size = 3) +  # Data points
  geom_line(data = predicted_data, aes(x = num_spec, y = connect, color = region), size = 1) +  # Fitted inverse lines
  labs(
    x = "Number of Species",
    y = "Connectance",
    color = "Region",
    title = ""
  ) +
  theme_minimal() +
  theme(legend.position = "top") 
b
```

The tropical region starts with a higher intercept (1.4789 higher on the log scale) than the temperate region. This suggests that at low species richness, tropical regions have more links than temperate regions. The slope for log number of species is lower in the tropical region by 0.5367 compared to the temperate region. This means that while tropical regions may have more links at lower species richness, the rate of increase in links is slower in tropical regions compared to temperate ones.

```{r assumptions_link}
# Residual plots to check assumptions
par(mfrow=c(2,2))
plot(mod_links)

# Check normality of residuals
hist(residuals(mod_links), main = "Histogram of Residuals", xlab = "Residuals", col = "lightblue")
qqnorm(residuals(mod_links))
qqline(residuals(mod_links))

# Extract residuals
residuals_links <- residuals(mod_links)
# Perform Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals_links)
shapiro_test

# high vif but that's exacly what we're trying to show 
vif(mod_links)
```

Assumptions:

1 Linearity

-   The relationship between species richness and connectance is linear within each region

2 **Homoscedasticity** (equal variance of residuals): Plot residuals vs fitted values to check for patterns

-   no clear pattern

**3 Normality of Residuals** (residuals should be normally distributed)

-   QQ plot looks normal

-   Shapiro-Wilk returns a high p-value

4 Independence

-   each stream represented once to reduce spatial autocorrelation

```{r assumptions_connect}
# Residual plots to check assumptions
par(mfrow=c(2,2))
plot(mod_connect)

# Check normality of residuals
hist(residuals(mod_connect), main = "Histogram of Residuals", xlab = "Residuals", col = "lightblue")
qqnorm(residuals(mod_connect))
qqline(residuals(mod_connect))

# Extract residuals
residuals_ancova <- residuals(mod_connect)
# Perform Shapiro-Wilk test for normality
shapiro_test <- shapiro.test(residuals_ancova)
shapiro_test

# high vif but that's exacly what we're trying to show 
vif(mod_connect)
```

Assumptions:

1 Linearity

-   The relationship between species richness and connectance is linear within each region

2 **Homoscedasticity** (equal variance of residuals): Plot residuals vs fitted values to check for patterns

-   no clear pattern, but skew left

**3 Normality of Residuals** (residuals should be normally distributed)

-   QQ plot looks normal

-   Shapiro-Wilk returns a high p-value

4 Independence

-   each stream represented once to reduce spatial autocorrelation
