---
title: "Exploring Cheddar"
output: html_document
date: "2024-08-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the tidyverse package
pacman::p_load(tidyverse)

library(readxl)
library(cheddar)

# Load in stream data 
contenido <- read_excel("contenido.xlsx")
```

## Exploring Cheddar

```{r sampledata}
data(TL84)
nps <- NPS(TL84) #dataframe of 
summary

NumberOfNodes(TL84) # equivalent to # taxon
NumberOfNodesByClass(TL84)
FractionOfNodesByClass(TL84)

NumberOfTrophicLinks(TL84)
LinkageDensity(TL84)
DirectedConnectance(TL84)

PlotWebByLevel(TL84)
#PlotWagonWheel(TL84)
```

## Urban Stream Data 
Using the stream data collected in an urban area, we will be looking at the:

- Number of Species (S) NumberOfNodesByClass() ?
- Number of Links (L) NumberOfTrophicLinks()
- Linkage Density (L/S) 
- Connectance (C=L/S^2)

okay this is not working, but i need to read in each sheet and then make a data frame, then turn it into a community object (?)

So each node is one species 
one big all inclusive food web, a second food web that excludes links that have less than 5% in the diet (or something) because some matter can come from the prey. for instance, a dragonfly is a predator but sometimes when they catch their prey they also catch algae that come with inverebrates that they are eating. This 5% number is just made up, its not really based on anything lol(trichotera are herbivores so they will get algae from eating these organisms... can pass into the stomach and be included in the food intake even tho they are not actually eating algae)

```{r practice}
dir <- "/Users/reillyconnelly/Desktop/THESIS"
# Manually read the nodes.csv file
nodes_data <- read.csv("nodes.csv", header = TRUE, stringsAsFactors = FALSE) #species column must be named "node'
links_data <- read.csv("links.csv", header = TRUE, stringsAsFactors = FALSE) #columns must be named 'resource' and 'consumer'

food_web <- Community(
  properties = list(
    title = "My Food Web",
    location = "Costa Rica",     # Overall community location
    ecosystem = "Tropical Rainforest",
    sampling_year = 2023),
  nodes = nodes_data ,# Provide nodes directly
  trophic.links = links_data          # Provide trophic links directly
)



# Check if the community loaded correctly
summary(food_web)

#YIPPEE

NumberOfNodes(food_web) # equivalent to # taxon
NumberOfNodesByClass(food_web)
FractionOfNodesByClass(food_web)

NumberOfTrophicLinks(food_web)
LinkageDensity(food_web)
DirectedConnectance(food_web)

PlotWebByLevel(food_web)
```
For the nodes data, I will be just taking the average of the mm and putting it in a new excel sheet 
For the links, I am including each food item that was observed in the stomach (excluding 'other'). Later on, I will make a new analysis removing food items that were not significant in the diet

```{r urbanstream}
dir <- "/Users/reillyconnelly/Desktop/THESIS"
# Manually read the nodes.csv file
urban_nodes <- read.csv("urban_nodes.csv", header = TRUE, stringsAsFactors = FALSE) 
urban_links <- read.csv("urban_links.csv", header = TRUE, stringsAsFactors = FALSE)

urban_web <- Community(
  properties = list(
    title = "Urban Stream Food Web",
    location = "Costa Rica",     # Overall community location
    ecosystem = "Tropical Rainforest",
    sampling_year = 2023),
  nodes = urban_nodes ,# Provide nodes directly
  trophic.links = urban_links          # Provide trophic links directly
)



# Check if the community loaded correctly
summary(urban_web)

#YIPPEE

NumberOfNodes(urban_web) # equivalent to # taxon
NumberOfNodesByClass(urban_web)
FractionOfNodesByClass(urban_web)

NumberOfTrophicLinks(urban_web)
LinkageDensity(urban_web)
DirectedConnectance(urban_web)

PlotWebByLevel(urban_web)
```

