---
title: "mixed model"
output: html_document
date: "2025-04-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(ggeffects)
library(readxl)
library(lme4)
library(lmerTest) # adds p values to summary 
data <- read_xlsx('data.xlsx', sheet='final_data')

# Exclude the last row (Yule et al. 2010) because data includes algae as species richness which influences the metrics
# Results significant with and without it 
data <- data[-nrow(data), ]
data <- data %>%
  mutate(log_num_spec = log(num_spec),
         log_num_link = log(num_link), 
         log_connect = log(connect))
head(data)
```

## Linear Mixed Model

I used a linear mixed effects model here to examine whether food web structure differed by region while controlling for the variation between studies and different sampling methods (paper as a random effect)

```{r , echo = FALSE}
lme.mod <- lmer(log_num_link ~ log_num_spec * region + (1 | paper), data = data)
summary(lme.mod)
```

The p-value 0.00888 for the interaction term, log_num_spec:regiontropical, in the log links model is significant, which suggests that the scaling of links with species differs between regions.

log temperate links slope = 1.5766

log tropical links slope = 1.5766 - 0.3074 = 1.2692

```{r}
# Get predictions from model
preds <- ggpredict(lme.mod, terms = c("log_num_spec", "region"))

ggplot(data, aes(x = log_num_spec, y = log_num_link, 
                      color = region)) +
  geom_point(alpha = 0.6, size = 2) +  # OG data points

  geom_line(data = preds, aes(x = x, y = predicted, color = group), size = 1.2) +

  labs(
    title = "Predicted Log(Number of Links) by Region",
    x = "Log(Number of Species)",
    y = "Log(Number of Links)",
    color = "Region",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(legend.position = "top")
```

This graph shows the predicted log number of link values (slopes and intercepts) from the mixed effects model, lme.mod, averaging over the random effect (paper)

```{r}
lme.mod2 <- lmer(log_connect ~ log_num_spec * region + (1 | paper), data = data)
summary(lme.mod2)
```

The p-value 0.00911 for the interaction term, log_num_spec:regiontropical, in the log connectance model is significant, which suggests that the scaling of connectance with species differs between regions.

log temperate connectance slope = -0.42459

log tropical links slope = -0.42459 - 0.30699 = -0.73158

```{r}
# Get predictions from model
preds2 <- ggpredict(lme.mod2, terms = c("log_num_spec", "region"))
ggplot(data, aes(x = log_num_spec, y = log_connect, 
                      color = region)) +
  geom_point(alpha = 0.6, size = 2) +  # OG data points

  geom_line(data = preds2, aes(x = x, y = predicted, color = group), size = 1.2) +

  labs(
    title = "Predicted Log(Connectance) by Region",
    x = "Log(Number of Species)",
    y = "Log(Connectance)",
    color = "Region",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(legend.position = "top")
```

This graph shows the predicted log connectance values (slopes and intercepts) from the mixed effects model, lme.mod2, averaging over the random effect (paper)

## Diagnostics
```{r}
plot(lme.mod)  # residuals vs fitted
qqnorm(resid(lme.mod)); qqline(resid(lme.mod))  # normality check, tails deviate a bit implying heavy tails i think

plot(lme.mod2)  
qqnorm(resid(lme.mod2)); qqline(resid(lme.mod2)) # tails deviate a bit 
```

